Simulating Ultra High-Dimensional Multivariate Data Using the bigsimr R Package

### 2020-06-15 docs/article_bigsimr.tex
migrate to html based preview

@@

\usepackage{graphicx,grffile}

@@

\title{Simulating Ultra High-Dimensional Multivariate Data Using the bigsimr R Package}
\author{A.G. Schissler \and A. Knudson}
\date{Thursday, June 04, 2020}

@@

\begin{verbatim}
       [,1]     [,2] [,3]
 [1,] 3.046 0.003646    2
 [2,] 3.180 0.091149    5
 [3,] 3.070 0.014344    2
 [4,] 3.157 0.164819    2
 [5,] 3.222 0.529193    5
 [6,] 3.118 0.341439    5
 [7,] 3.056 0.028928    2
 [8,] 3.187 0.119534    4
 [9,] 3.118 0.413526    3
[10,] 3.225 0.431735    3
\end{verbatim}

@@

\begin{verbatim}
       [,1]   [,2]   [,3]
[1,] 1.0000 0.4886 0.4945
[2,] 0.4886 1.0000 0.4882
[3,] 0.4945 0.4882 1.0000
\end{verbatim}

We can see that even with 10,000 samples, the estimated correlation of the simulated data is not exactly the same as the target correlation. This can be explained by the fact that some correlations are simply not possible due to the discrete nature of certain distributions. Another possibility is that the copula algorithm is biased and needs correction.

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{## computation times normal}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{## full workflow simulating RNA-seq data}
\end{Highlighting}
\end{Shaded}

@@

@@

\begin{itemize}
\tightlist
\item
  Kendall's \(\tau\) was estimated efficiently using the \texttt{pcaPP::cor.fk} function in \texttt{R}.
\end{itemize}

\hypertarget{ultra-high-dimensional-simulation-based-correlation-hypothesis-testing}{%
\subsection{Ultra High-Dimensional simulation-based correlation hypothesis testing}\label{ultra-high-dimensional-simulation-based-correlation-hypothesis-testing}}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{## full workflow simulating RNA-seq data}
\end{Highlighting}
\end{Shaded}

@@

\CommentTok{## simulation based}
\CommentTok{## devtools::install_github("adknudson/bigsimr", ref="develop")}
\KeywordTok{library}\NormalTok{(bigsimr)}
\CommentTok{## ?rvec}

\NormalTok{d <-}\StringTok{ }\DecValTok{5}
\NormalTok{rho <-}\StringTok{ }\KeywordTok{rcor}\NormalTok{(d)}
\NormalTok{margins <-}\StringTok{ }\KeywordTok{list}\NormalTok{(}
    \KeywordTok{list}\NormalTok{(}\StringTok{"nbinom"}\NormalTok{, }\DataTypeTok{size =} \DecValTok{5}\NormalTok{, }\DataTypeTok{prob =} \FloatTok{0.3}\NormalTok{),}
    \KeywordTok{list}\NormalTok{(}\StringTok{"exp"}\NormalTok{, }\DataTypeTok{rate =} \DecValTok{4}\NormalTok{),}
    \KeywordTok{list}\NormalTok{(}\StringTok{"binom"}\NormalTok{, }\DataTypeTok{size =} \DecValTok{5}\NormalTok{, }\DataTypeTok{prob =} \FloatTok{0.7}\NormalTok{),}
    \KeywordTok{list}\NormalTok{(}\StringTok{"norm"}\NormalTok{, }\DataTypeTok{mean =} \DecValTok{10}\NormalTok{, }\DataTypeTok{sd =} \DecValTok{3}\NormalTok{),}
    \KeywordTok{list}\NormalTok{(}\StringTok{"pois"}\NormalTok{, }\DataTypeTok{lambda =} \DecValTok{10}\NormalTok{)}
\NormalTok{)}

\NormalTok{params =}\StringTok{ }\NormalTok{margins}
\NormalTok{rho =}\StringTok{ }\NormalTok{rho}
\CommentTok{## cores = parallel::detectCores() - 1}
\NormalTok{cores =}\StringTok{ }\DecValTok{2}
\NormalTok{reps =}\StringTok{ }\FloatTok{1e3}
\NormalTok{type =}\StringTok{ "spearman"}

\CommentTok{## help(package="bigsimr")}
\NormalTok{rho_bounds <-}\StringTok{ }\KeywordTok{computeCorBounds}\NormalTok{(}\DataTypeTok{params =}\NormalTok{ margins, }\DataTypeTok{cores =}\NormalTok{ cores, }\DataTypeTok{type =}\NormalTok{ type, }\DataTypeTok{reps =}\NormalTok{ reps)}
\NormalTok{rho_bounds}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{## full workflow simulating RNA-seq data}
\end{Highlighting}
\end{Shaded}

@@

>>>>

\usepackage[margin=1in]{geometry}

@@

\usepackage{graphicx}

@@

\title{Simulating Ultra High-Dimensional Multivariate Data Using the bigsimr R Package}
\author{A.G. Schissler \and A. Knudson}
\date{Monday, June 15, 2020}

@@

\begin{verbatim}
        [,1]     [,2] [,3]
  [1,] 3.044 0.095446    3
  [2,] 3.266 0.149347    4
  [3,] 3.210 0.368758    7
  [4,] 3.096 0.273881    1
  [5,] 3.262 0.167132    8
  [6,] 3.186 0.297752    4
  [7,] 3.244 0.378209    2
  [8,] 3.092 0.061396    2
  [9,] 3.153 0.388307    5
 [10,] 3.191 0.017892    2
 [11,] 3.021 0.143984    1
 [12,] 2.977 0.306241    2
 [13,] 3.327 0.311243    6
 [14,] 3.087 0.166009    3
 [15,] 3.146 0.462549    2
 [16,] 3.130 0.045254    3
 [17,] 3.215 0.341481    3
 [18,] 3.153 0.481414    9
 [19,] 3.311 0.536290    6
 [20,] 3.054 0.183528    4
 [21,] 3.017 0.011316    1
 [22,] 3.195 0.293363    1
 [23,] 3.017 0.077518    1
 [24,] 3.024 0.041316    1
 [25,] 3.272 0.330607    6
 [26,] 3.218 0.206071    5
 [27,] 2.984 0.001637    2
 [28,] 3.156 0.114774    3
 [29,] 3.269 0.089939    3
 [30,] 3.103 0.177539    3
 [31,] 3.147 0.138347    6
 [32,] 3.157 0.093564    4
 [33,] 3.318 0.356667    2
 [34,] 3.111 0.231386    4
 [35,] 3.174 0.659264    2
 [36,] 3.218 0.293820    4
 [37,] 3.264 0.240420    6
 [38,] 3.072 0.255088    3
 [39,] 3.131 0.168695    0
 [40,] 3.280 0.491158    6
 [41,] 3.189 0.238835    5
 [42,] 3.017 0.011214    1
 [43,] 3.145 0.022872    1
 [44,] 3.111 0.144358    1
 [45,] 3.236 0.085572    3
 [46,] 3.221 0.309900    7
 [47,] 2.955 0.014524    3
 [48,] 3.069 0.011569    2
 [49,] 2.970 0.075424    2
 [50,] 3.265 0.233258    5
 [51,] 3.125 0.223936    2
 [52,] 3.118 0.100138    3
 [53,] 3.042 0.113373    2
 [54,] 2.993 0.335292    3
 [55,] 3.073 0.020944    1
 [56,] 3.178 0.256723    4
 [57,] 3.242 0.679650    7
 [58,] 3.204 0.228560    5
 [59,] 3.325 0.254273    9
 [60,] 2.918 0.109870    2
 [61,] 3.106 0.134519    5
 [62,] 3.035 0.022232    1
 [63,] 3.238 0.029679    3
 [64,] 3.163 0.159671    3
 [65,] 3.250 0.265043    5
 [66,] 3.139 0.109471    6
 [67,] 2.952 0.069780    3
 [68,] 3.092 0.103947    6
 [69,] 3.163 0.196133    6
 [70,] 2.993 0.079026    1
 [71,] 2.990 0.043229    2
 [72,] 3.057 0.018904    2
 [73,] 2.985 0.009260    1
 [74,] 3.224 0.778705    6
 [75,] 3.231 0.238377    4
 [76,] 3.083 0.014756    2
 [77,] 3.175 0.262330    1
 [78,] 3.141 0.106700    3
 [79,] 3.047 0.007255    2
 [80,] 3.182 0.334552    2
 [81,] 3.184 0.303953    5
 [82,] 3.033 0.182557    2
 [83,] 3.114 0.227299    4
 [84,] 3.066 0.057184    5
 [85,] 3.077 0.069102    4
 [86,] 2.866 0.004398    1
 [87,] 2.920 0.100838    2
 [88,] 3.219 0.175613    6
 [89,] 3.080 0.094636    1
 [90,] 3.145 0.076677    4
 [91,] 3.176 0.243460    4
 [92,] 3.005 0.230508    2
 [93,] 3.056 0.315420    5
 [94,] 3.162 0.524840    5
 [95,] 3.332 0.685218    7
 [96,] 3.120 0.209939    0
 [97,] 3.184 0.174008    2
 [98,] 3.037 0.037463    2
 [99,] 3.064 0.013679    1
[100,] 3.098 0.101754    3
\end{verbatim}

@@

\begin{verbatim}
       [,1]   [,2]   [,3]
[1,] 1.0000 0.4989 0.4963
[2,] 0.4989 1.0000 0.4929
[3,] 0.4963 0.4929 1.0000
\end{verbatim}

We can see that even with 100,000 samples, the estimated correlation of the simulated data is not exactly the same as the target correlation. This can be explained by the fact that some correlations are simply not possible due to the discrete nature of certain distributions. Another possibility is that the copula algorithm is biased and needs correction.

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#\# computation times normal}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#\# full workflow simulating RNA{-}seq data}
\NormalTok{allDat \textless{}{-}}\StringTok{ }\KeywordTok{readRDS}\NormalTok{( }\DataTypeTok{file =} \StringTok{"\textasciitilde{}/Downloads/complete\_processed\_tcga2stat\_RNASeq2\_with\_clinical.rds"}\NormalTok{ )}
\NormalTok{lastClinical \textless{}{-}}\StringTok{ }\KeywordTok{which}\NormalTok{( }\KeywordTok{names}\NormalTok{(allDat) }\OperatorTok{==}\StringTok{ \textquotesingle{}tumorsize\textquotesingle{}}\NormalTok{ )}
\NormalTok{brca \textless{}{-}}\StringTok{ }\NormalTok{allDat[ allDat}\OperatorTok{$}\NormalTok{disease }\OperatorTok{==}\StringTok{ "BRCA"}\NormalTok{, (lastClinical}\OperatorTok{+}\DecValTok{1}\NormalTok{)}\OperatorTok{:}\KeywordTok{ncol}\NormalTok{(allDat) ]}
\CommentTok{\#\# remove naively the RSEM adjustment}
\NormalTok{brca \textless{}{-}}\StringTok{ }\KeywordTok{round}\NormalTok{(brca, }\DecValTok{0}\NormalTok{)}
\KeywordTok{ncol}\NormalTok{(brca) }\CommentTok{\#\# num of genes 20501}
\NormalTok{[}\DecValTok{1}\NormalTok{] }\DecValTok{20501}
\CommentTok{\#\# compute the median expression! avoid the 0s}
\NormalTok{brcaMedian \textless{}{-}}\StringTok{ }\KeywordTok{apply}\NormalTok{(brca, }\DecValTok{2}\NormalTok{, median)}
\CommentTok{\#\# retain top (1{-}probs)*100\% highest expressing genes for illustration}
\NormalTok{myProb \textless{}{-}}\StringTok{ }\FloatTok{0.95} 
\NormalTok{cutPoint \textless{}{-}}\StringTok{ }\KeywordTok{quantile}\NormalTok{( }\DataTypeTok{x =}\NormalTok{ brcaMedian, }\DataTypeTok{probs =}\NormalTok{ myProb )}
\CommentTok{\#\# genesToKeep \textless{}{-} names( brcaMedian ) [ which(brcaMedian \textgreater{}= cutPoint) ]}
\NormalTok{genesToKeep \textless{}{-}}\StringTok{ }\KeywordTok{names}\NormalTok{( brcaMedian ) [ }\KeywordTok{which}\NormalTok{(brcaMedian }\OperatorTok{\textgreater{}=}\StringTok{ }\NormalTok{cutPoint) ]}
\NormalTok{brca \textless{}{-}}\StringTok{ }\NormalTok{brca[ , genesToKeep ]}
\CommentTok{\#\# ncol(brca) / 20501}
\NormalTok{(d \textless{}{-}}\StringTok{ }\KeywordTok{length}\NormalTok{(genesToKeep))}
\NormalTok{[}\DecValTok{1}\NormalTok{] }\DecValTok{1026}
\end{Highlighting}
\end{Shaded}

@@

\NormalTok{logVarOverMean \textless{}{-}}\StringTok{ }\ControlFlowTok{function}\NormalTok{( x ) \{}
    \KeywordTok{log}\NormalTok{ ( }\KeywordTok{var}\NormalTok{(x) }\OperatorTok{/}\StringTok{ }\KeywordTok{mean}\NormalTok{ (x) )}
\NormalTok{\}}
\CommentTok{\#\# check whether NB makes sense}

\NormalTok{summaryBRCA \textless{}{-}}\StringTok{ }\NormalTok{brca }\OperatorTok{\%\textgreater{}\%}\StringTok{ }
\StringTok{    }\KeywordTok{summarize\_all}\NormalTok{(  }\KeywordTok{list}\NormalTok{(}\OperatorTok{\textasciitilde{}}\StringTok{ }\KeywordTok{logVarOverMean}\NormalTok{( . ) ) )}
\NormalTok{summaryBRCA \textless{}{-}}\StringTok{ }\KeywordTok{as.data.frame}\NormalTok{( }\KeywordTok{t}\NormalTok{(}\KeywordTok{as.data.frame}\NormalTok{( summaryBRCA )) )}
\KeywordTok{names}\NormalTok{(summaryBRCA) \textless{}{-}}\StringTok{ "logVarOverMean"}
\CommentTok{\#\# sum(summaryBRCA$logVarOverMean \textless{} 0)  \#\# no underdispersed genes}
\KeywordTok{ggplot}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ summaryBRCA, }\DataTypeTok{mapping =} \KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ logVarOverMean)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom\_histogram}\NormalTok{(}\DataTypeTok{color =} \StringTok{"white"}\NormalTok{, }\DataTypeTok{bins =} \DecValTok{20}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}

{\centering \includegraphics[width=0.75\linewidth]{article_bigsimr_files/figure-latex/figNBratio-1}

\caption{The genes display extreme overdispersion. Using group-specific sample means and variances for each of the d=`r d` genes, we display the distributions oflog of variance to mean ratio.}\label{fig:figNBratio}
\end{figure}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#\# full workflow simulating RNA{-}seq data}

\CommentTok{\#\# 1. Estimate Spearman\textquotesingle{}s correlation on the count data}
\NormalTok{corType \textless{}{-}}\StringTok{ \textquotesingle{}spearman\textquotesingle{}}
\KeywordTok{system.time}\NormalTok{( rho \textless{}{-}}\StringTok{ }\NormalTok{bigsimr}\OperatorTok{::}\KeywordTok{fastCor}\NormalTok{( brca, }\DataTypeTok{method =}\NormalTok{ corType ) )}
\NormalTok{   user  system elapsed }
  \FloatTok{0.654}   \FloatTok{0.013}   \FloatTok{0.673} 
\CommentTok{\#\# Describe the correlations}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#\# 2. Estimate NegBin parameters using Method of Moments}
\NormalTok{estimateNegBinMoM \textless{}{-}}\StringTok{ }\ControlFlowTok{function}\NormalTok{(tmpGene, }\DataTypeTok{minP =} \FloatTok{1e{-}5}\NormalTok{) \{}
\NormalTok{    tmpMean \textless{}{-}}\StringTok{ }\KeywordTok{mean}\NormalTok{(tmpGene)}
\NormalTok{    tmpVar \textless{}{-}}\StringTok{ }\KeywordTok{var}\NormalTok{(tmpGene)}
    \CommentTok{\#\# relate to nbinom parameters}
    \CommentTok{\#\# See ?rbinom for details.}
\NormalTok{    p \textless{}{-}}\StringTok{ }\NormalTok{tmpMean }\OperatorTok{/}\StringTok{ }\NormalTok{tmpVar}
    \ControlFlowTok{if}\NormalTok{ (p }\OperatorTok{\textless{}}\StringTok{ }\NormalTok{minP) \{p \textless{}{-}}\StringTok{ }\NormalTok{minP \} }\CommentTok{\#\# maybe add noise here}
\NormalTok{    n \textless{}{-}}\StringTok{ }\NormalTok{( tmpMean }\OperatorTok{*}\StringTok{ }\NormalTok{p) }\OperatorTok{/}\StringTok{ }\NormalTok{( }\DecValTok{1}  \OperatorTok{{-}}\StringTok{ }\NormalTok{p )  }
    \CommentTok{\#\# format for bigsimr margins}
    \KeywordTok{return}\NormalTok{( }\KeywordTok{list}\NormalTok{(}\StringTok{"nbinom"}\NormalTok{, }\DataTypeTok{size =}\NormalTok{ n, }\DataTypeTok{prob =}\NormalTok{ p) )}
\NormalTok{\}}
\NormalTok{brcaMargins \textless{}{-}}\StringTok{ }\KeywordTok{apply}\NormalTok{( }\KeywordTok{unname}\NormalTok{(}\KeywordTok{as.matrix}\NormalTok{(brca)), }\DecValTok{2}\NormalTok{, estimateNegBinMoM )}
\CommentTok{\#\# describe the margins}
\CommentTok{\#\# check for too small prob}
\CommentTok{\#\# head( sort( unlist( lapply( brcaMargins, function(x) \{x$prob\} ) ) ) )}
\CommentTok{\#\# head( sort( unlist( lapply( brcaMargins, function(x) \{x$size\} ) ) ) )}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#\# 3. Generate the simulated samples}
\CommentTok{\#\# N \textless{}{-} 10}
\CommentTok{\#\# system.time( simBRCA \textless{}{-} rvec(N, rho = rho[1:2, 1:2], params = brcaMargins[1:2], type = corType, adjustForDiscrete = FALSE) )}
\NormalTok{N \textless{}{-}}\StringTok{ }\KeywordTok{nrow}\NormalTok{(brca)}
\KeywordTok{system.time}\NormalTok{( simBRCA \textless{}{-}}\StringTok{ }\KeywordTok{rvec}\NormalTok{(N, }\DataTypeTok{rho =}\NormalTok{ rho, }\DataTypeTok{params =}\NormalTok{ brcaMargins, }\DataTypeTok{cores =}\NormalTok{ CORES, }\DataTypeTok{type =}\NormalTok{ corType, }\DataTypeTok{adjustForDiscrete =} \OtherTok{FALSE}\NormalTok{) )}
\NormalTok{simBRCA[}\DecValTok{1}\OperatorTok{:}\DecValTok{2}\NormalTok{, }\DecValTok{1}\OperatorTok{:}\DecValTok{2}\NormalTok{ ]}
\end{Highlighting}
\end{Shaded}

\CommentTok{\#\# Describe and plot real data}
\NormalTok{GGally}\OperatorTok{::}\KeywordTok{ggpairs}\NormalTok{(}\DataTypeTok{data =} \KeywordTok{as.data.frame}\NormalTok{(brca[ ,}\DecValTok{1}\OperatorTok{:}\DecValTok{3}\NormalTok{] ) )}

\CommentTok{\#\# Describe and plot simulations}
\NormalTok{GGally}\OperatorTok{::}\KeywordTok{ggpairs}\NormalTok{(}\DataTypeTok{data =} \KeywordTok{as.data.frame}\NormalTok{(simBRCA[ ,}\DecValTok{1}\OperatorTok{:}\DecValTok{3}\NormalTok{] ) )}

\CommentTok{\#\# check 1st moments}
\NormalTok{trueMu \textless{}{-}}\StringTok{ }\KeywordTok{colMeans}\NormalTok{(brca)}
\NormalTok{simMu \textless{}{-}}\StringTok{ }\KeywordTok{colMeans}\NormalTok{(simBRCA)}
\KeywordTok{qplot}\NormalTok{( }\DataTypeTok{x =}\NormalTok{ trueMu, }\DataTypeTok{y =}\NormalTok{ simMu ) }\OperatorTok{+}\StringTok{ }\KeywordTok{geom\_abline}\NormalTok{(}\DataTypeTok{slope =} \DecValTok{1}\NormalTok{, }\DataTypeTok{intercept =} \DecValTok{0}\NormalTok{)}

\CommentTok{\#\# check 2nd moment}
\NormalTok{trueVar \textless{}{-}}\StringTok{ }\KeywordTok{apply}\NormalTok{( brca, }\DecValTok{2}\NormalTok{, var)}
\NormalTok{simVar \textless{}{-}}\StringTok{ }\KeywordTok{apply}\NormalTok{( simBRCA, }\DecValTok{2}\NormalTok{, var)}
\KeywordTok{qplot}\NormalTok{( }\DataTypeTok{x =}\NormalTok{ trueVar, }\DataTypeTok{y =}\NormalTok{ simVar ) }\OperatorTok{+}\StringTok{ }\KeywordTok{geom\_abline}\NormalTok{(}\DataTypeTok{slope =} \DecValTok{1}\NormalTok{, }\DataTypeTok{intercept =} \DecValTok{0}\NormalTok{)}

\CommentTok{\#\# check correlation}
\KeywordTok{system.time}\NormalTok{( simRho \textless{}{-}}\StringTok{ }\NormalTok{bigsimr}\OperatorTok{::}\KeywordTok{fastCor}\NormalTok{( simBRCA, }\DataTypeTok{method =}\NormalTok{ corType ) )}
\NormalTok{rho[}\DecValTok{1}\OperatorTok{:}\DecValTok{5}\NormalTok{, }\DecValTok{1}\OperatorTok{:}\DecValTok{5}\NormalTok{]}
\NormalTok{simRho[}\DecValTok{1}\OperatorTok{:}\DecValTok{5}\NormalTok{, }\DecValTok{1}\OperatorTok{:}\DecValTok{5}\NormalTok{]}
\NormalTok{trueRho \textless{}{-}}\StringTok{ }\NormalTok{rho[}\KeywordTok{lower.tri}\NormalTok{(rho)]}
\NormalTok{simRho \textless{}{-}}\StringTok{ }\NormalTok{simRho[}\KeywordTok{lower.tri}\NormalTok{(simRho)]}
\KeywordTok{qplot}\NormalTok{( }\DataTypeTok{x =}\NormalTok{ trueRho, }\DataTypeTok{y =}\NormalTok{ simRho ) }\OperatorTok{+}\StringTok{ }\KeywordTok{geom\_abline}\NormalTok{(}\DataTypeTok{slope =} \DecValTok{1}\NormalTok{, }\DataTypeTok{intercept =} \DecValTok{0}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{ultra-high-dimensional-simulation-based-correlation-hypothesis-testing-under}{%
\subsection{Ultra High-Dimensional simulation-based correlation hypothesis testing under}\label{ultra-high-dimensional-simulation-based-correlation-hypothesis-testing-under}}

Since many of the correlations are near zero, one could ask whether it is
possible that

@@

\CommentTok{\#\# simulation based}
\CommentTok{\#\# devtools::install\_github("adknudson/bigsimr", ref="develop")}
\KeywordTok{library}\NormalTok{(bigsimr)}
\CommentTok{\#\# ?rvec}

\NormalTok{d \textless{}{-}}\StringTok{ }\DecValTok{5}
\NormalTok{rho \textless{}{-}}\StringTok{ }\KeywordTok{rcor}\NormalTok{(d)}
\NormalTok{margins \textless{}{-}}\StringTok{ }\KeywordTok{list}\NormalTok{(}
    \KeywordTok{list}\NormalTok{(}\StringTok{"nbinom"}\NormalTok{, }\DataTypeTok{size =} \DecValTok{5}\NormalTok{, }\DataTypeTok{prob =} \FloatTok{0.3}\NormalTok{),}
    \KeywordTok{list}\NormalTok{(}\StringTok{"exp"}\NormalTok{, }\DataTypeTok{rate =} \DecValTok{4}\NormalTok{),}
    \KeywordTok{list}\NormalTok{(}\StringTok{"binom"}\NormalTok{, }\DataTypeTok{size =} \DecValTok{5}\NormalTok{, }\DataTypeTok{prob =} \FloatTok{0.7}\NormalTok{),}
    \KeywordTok{list}\NormalTok{(}\StringTok{"norm"}\NormalTok{, }\DataTypeTok{mean =} \DecValTok{10}\NormalTok{, }\DataTypeTok{sd =} \DecValTok{3}\NormalTok{),}
    \KeywordTok{list}\NormalTok{(}\StringTok{"pois"}\NormalTok{, }\DataTypeTok{lambda =} \DecValTok{10}\NormalTok{)}
\NormalTok{)}

\NormalTok{params =}\StringTok{ }\NormalTok{margins}
\NormalTok{rho =}\StringTok{ }\NormalTok{rho}
\CommentTok{\#\# cores = parallel::detectCores() {-} 1}
\NormalTok{cores =}\StringTok{ }\DecValTok{2}
\NormalTok{reps =}\StringTok{ }\FloatTok{1e3}
\NormalTok{type =}\StringTok{ "spearman"}

\CommentTok{\#\# help(package="bigsimr")}
\NormalTok{rho\_bounds \textless{}{-}}\StringTok{ }\KeywordTok{computeCorBounds}\NormalTok{(}\DataTypeTok{params =}\NormalTok{ margins, }\DataTypeTok{cores =}\NormalTok{ cores, }\DataTypeTok{type =}\NormalTok{ type, }\DataTypeTok{reps =}\NormalTok{ reps)}
\NormalTok{rho\_bounds}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#\# full workflow simulating RNA{-}seq data}
\end{Highlighting}
\end{Shaded}

@@

Mention \texttt{rslurm}

### 2020-06-15 index.Rmd
migrate to html based preview

--- 
title: "Simulating Ultra High-Dimensional Multivariate Data Using the bigsimr R Package"
author: [A.G. Schissler, A. Knudson]
date: "`r format(Sys.Date(), '%A, %B %d, %Y')`"
site: bookdown::bookdown_site
output: 
  bookdown::gitbook:
    base_format: rticles::tf_article
documentclass: article
bibliography: [bigsimr.bib, packages.bib]
biblio-style: apalike
link-citations: yes
abstract: |
  In this era of Big Data, it is critical to realistically simulate data to conduct informative Monte Carlo studies. This is problematic when data are inherently multivariate while at the same time are (ultra-) high dimensional. This situation appears frequently in observational data found on online and in high-throughput biomedical experiments (e.g., RNA-sequencing). Due to the difficulty in simulating realistic correlated data points, researchers often resort to simulation designs that posit independence --- greatly diminishing the insight into the empirical operating characteristics of any proposed methodology.  Major challenges lie in the computational complexity involved in simulating these massive random vectors. We propose a fairly general, scalable procedure to simulate high-dimensional multivariate distributions with pre-specified marginal characteristics and dependency characteristics. As a motivating example, we use our methodology to study large-scale statistical inferential procedures applied to cancer-related RNA-sequencing data sets. The proposed algorithm is implemented as the `bigsimr` R package.
---

```{r LoadLib, echo = FALSE, eval = TRUE, message=FALSE, warning=FALSE, results = "hide"}
library(ggplot2)
library(tidyverse)
library(knitr)
library(dplyr)
set.seed(06082020)
```

>>>>

--- 
title: "Simulating Ultra High-Dimensional Multivariate Data Using the bigsimr R Package"
author: [A.G. Schissler, A. Knudson]
date: "`r format(Sys.Date(), '%A, %B %d, %Y')`"
site: bookdown::bookdown_site
bibliography: [bigsimr.bib, packages.bib]
biblio-style: apalike
link-citations: yes
abstract: |
  In this era of Big Data, it is critical to realistically simulate data to conduct informative Monte Carlo studies. This is problematic when data are inherently multivariate while at the same time are (ultra-) high dimensional. This situation appears frequently in observational data found on online and in high-throughput biomedical experiments (e.g., RNA-sequencing). Due to the difficulty in simulating realistic correlated data points, researchers often resort to simulation designs that posit independence --- greatly diminishing the insight into the empirical operating characteristics of any proposed methodology.  Major challenges lie in the computational complexity involved in simulating these massive random vectors. We propose a fairly general, scalable procedure to simulate high-dimensional multivariate distributions with pre-specified marginal characteristics and dependency characteristics. As a motivating example, we use our methodology to study large-scale statistical inferential procedures applied to cancer-related RNA-sequencing data sets. The proposed algorithm is implemented as the `bigsimr` R package.
---

```{r LoadLib, echo = FALSE, eval = TRUE, message=FALSE, warning=FALSE, results = "hide"}
library(ggplot2)
library(tidyverse)
library(knitr)
library(dplyr)
## devtools::install_github( "adknudson/bigsimr", ref = 'develop')
set.seed(06082020)
```
